### 2007-09-20 pkg/paper/article.Rnw
Added article files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@5 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{article}
\usepackage{graphicx}
\begin{document}

This article aims to set out the technical information that is
necessary for producing precise specifications
of colour within a software environment.

This article is aimed at the developers of software for
statistical graphics, so it is not an exhaustive discussion of
colour.  However, sufficient background material is included 
to allow an understanding of the terms and concepts used in this
article.

It is hoped that the information will also be generally useful
for people working with colours as data values, including 
the generation of sets of colours for statistical plots.

\section{What's wrong with RGB?}

The most common way to specify colours in a software system
is as an RGB triplet.  In this system, a colour is specified 
as an amount of red, an amount of green, and an amount of blue,
where the amounts range from 0 (none) to 1 (as much as possible).
In some systems, the amounts are specified in the range 0 to 255
(because it is common to use 8 bits to store each colour component
and $2^8 = 256$), and it is not uncommon to write values within this 
range in hexadecimal notation so that values range from \#00 to
\#FF (F has the decimal value 15 and the second digit 
specifies a multiple of
16, so $15*16 + 15 = 255$).

For example, in HTML, the colour red is specified as \#FF0000.

\subsection{RGB is unintuitive}

This system for specifying colour has some appeal.  It appears 
straightforward to specify simple colours:  red is all red and no green
or blue, green and blue are similar, white is all three colours, and black is
none.

However, intuition rapidly runs out of steam in this system.  What is the
RGB specification for the colour brown?  What about pink or orange?

It is possible to create colour palettes for the RGB system and 
to select colours by eye, but this is not appropriate for situations
where it is necessary to select sets of colours.  For example, 
it might be useful to represent an ordinal variable as different
shades of brown.  Given the RGB specification for brown, how 
can we specify lighter shades of brown?

\subsection{RGB is not perceptually uniform}

An intuitive answer to the previous question would be to linearly
interpolate between the RGB values for brown and the RGB values
for white, but this only serves to highlight another problem with the
RGB system, which is that it is not a \emph{perceptually uniform} system.
This means that a unit step within RGB space does not produce a 
constant perceptual change in colour.
Figure \ref{figure:shadesofbrown} shows the linear interpolation in
RGB space compared to a similar interpolation in a perceptually uniform space.

<<echo=FALSE>>=
library(grid)
library(colorspace)

<<rgbnotuniform, echo=FALSE, fig=TRUE, height=3, include=FALSE>>=
# grid.newpage()

# RGB
N <- 100
grid.rect(x=1:N/N, width=1/N, y=.75, height=0.4,
          just=c("right", "centre"),
          gp=gpar(col=NA,
            fill=rgb(seq(165, 255, length=N),
              seq(42, 255, length=N),
              seq(42, 255, length=N), max=255)))

# HCL

# col2rgb("brown")
#       [,1]
# red    165
# green   42
# blue    42
# N <- 20

# > as(RGB(165/255, 42/255, 42/255), "polarLAB")
#             L        C        H
# [1,] 58.72249 41.92013 24.42661

cols <- coords(as(polarLAB(seq(58.7224, 99, length=N),
                           seq(41.92013, 0, length=N),
                           24.42661), "RGB"))

grid.rect(x=1:N/N, width=1/N, y=0.25, height=0.4,
          just=c("right", "centre"),
          gp=gpar(col=NA, fill=rgb(cols[,1], cols[,2], cols[,3])))

@ 
\begin{figure}
\includegraphics[width=\textwidth]{article-rgbnotuniform}
\caption{\label{figure:shadesofbrown}A linear interpolation
from brown to white in RGB space (top) and a perceptually uniform
 space (bottom).  Equal-sized horizontal steps result in non-equal-sized
perceptual changes in colour in the RGB interpolation (the perceived 
colour changes rapidly at the left end and more slowly at the 
right end).}
\end{figure}

\subsection{RGB is device-dependent}

A further problem with RGB colour specifications is that it is unclear
what ``red'', ``green'', and ``blue'' actually mean.

The colour of light is determined by its wave-length. A typical human can 
perceive light in the range 400nm to 700nm.  A coherent
light source (e.g., a laser) at the 400nm range appears blue and a coherent
light source at the 700nm range appears red.  Most light sources
(e.g., light bulbs, or the Sun)
are not coherent and emit light at many different wave lengths.
A light source can be characterised by its Spectral Power Distribution (SPD),
which is how much of each wavelength of light the light source emits.

On an LCD computer screen, 
each pixel is made up from three sub-pixels, which are
coloured red, green, and blue.  A full red colour is produced by emitting 
light at full power from just the red sub-pixel.  However, this red is not
a single wavelength.
Figure \ref{figure:SPD} shows the SPDs might look like 
for the red, green, and blue
sub-pixels of a typical computer monitor.

% source: http://white.stanford.edu/~brian/scielab/scielab1-1-1/

<<spd, echo=FALSE, fig=TRUE, include=FALSE>>=
SPD <- read.table("displaySPD.txt", skip=5,
    col.names=c("r", "g", "b"))
wavelengths <- scan("wavelength.txt", what=numeric(), skip=5)

plot(wavelengths, SPD$r, type="l", col="red", lwd=3,
     ylab="", xlab="wavelength", axes=FALSE)
lines(wavelengths, SPD$g, type="l", col="green", lwd=3)
lines(wavelengths, SPD$b, type="l", col="blue", lwd=3)
box()
axis(1)

@ 
\begin{figure}
\includegraphics[width=\textwidth]{article-spd}
\caption{\label{figure:SPD}SPD for red, green, and blue on
a typical computer monitor.}
\end{figure}

In order to know what light is emitted in response to an RGB colour 
specification like \#FF0000, it is necessary to know the SPDs for
the device that will be producing the light.
This can vary considerably between different devices such as computer
monitors and data projectors (beamers).

For less-than-full red, the SPD stays the same shape and just
less light is emitted (the result is a darker colour).
However, there is a further complication because the decrease is not linear.
Figure \ref{figure:gamma} shows the curves relating the amount of
light emitted as a function of the sizes of the RGB components
for a typical computer monitor.
These are commonly referred to as ``display gamma curves'' because
the curve is, for convenience, approximated by a power function
with a single exponent called gamma.  Although each curve is slightly 
different, it is common to use a single gamma value for a
particular device.  Figure \ref{figure:gamma} shows the power curve
for the standard gamma value of 2.2 as a dotted line.

% source: http://white.stanford.edu/~brian/scielab/scielab1-1-1/

<<gamma, echo=FALSE, fig=TRUE, include=FALSE>>=
gamma <- read.table("gamma.txt", skip=5)

plot(seq(0, 1, length=dim(gamma)[1]), gamma[,1],
     type="l", col="red", lwd=3,
     xlab="amount of primary", ylab="intensity")
lines(seq(0, 1, length=dim(gamma)[1]), gamma[,2],
     type="l", col="green", lwd=3)
lines(seq(0, 1, length=dim(gamma)[1]), gamma[,3],
     type="l", col="blue", lwd=3)
lines(seq(0, 1, length=dim(gamma)[1]),
     seq(0, 1, length=dim(gamma)[1])^2.2,
     type="l", lty="dotted", lwd=3)

@ 
\begin{figure}
\includegraphics[width=\textwidth]{article-gamma}
\caption{\label{figure:gamma}Gamma curves for red, green, and blue on
a typical computer monitor.}
\end{figure}

In order to know how much light is emitted in response
to an RGB colour specification, for example, \#990000, it is necessary
to know the gamma for the device that will be producing the light. 
This can also vary considerably between different devices such as computer
monitors and data projectors (beamers).

In summary, we have what we will call the \emph{RGB uncertainty principle}:
in RGB space, you don't know where you are, 
you don't know how fast you are moving, and 
you don't know which way to go.

It is simply not adequate for a software system to only allow 
the user to specify colours in an RGB system.

\section{Device-independence:  CIE XYZ}

Human perception of colours.  Standard observer.

Chromaticities.

\section{Perceptual uniformity:  CIE Lab}

\section{Intuitive colour choice:  HCL}

\section{sRGB}

RGB is still important because that is how you express colours
in some systems, ergo it is necessary to allow people to work
in a nice colourspace, but then convert to RGB.

\section{Other colour spaces}

CMYK for print.  Simple-minded RGB-CMY is easy, but wrong.
Sophisticated RGB-CMY is damn nigh impossible.

\end{document}