### 2017-09-24 pkg/colorspace/inst/hclconverter/html/appInfo.Rmd
Added hcl_converter, alpha version.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@206 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

## App Information

### Terms and Conditions

Why this app ..

### Terms of Use

It is mutually assumed and understood that uploading and converting images with
this app is done at your own risk.  The creator/maintainer of the
``colorspace`` package and/or the [hclwizard.org](http://hclwizard.org) website
assume no responsibility for the content of the uploaded images and a possible
misuse of the uploaded images.

Uploaded images are stored in the computers temporary directory for the
conversion process only and will be deleted immediately after the results are
shown on the front-end.  Neither the uploaded image files or any of the results
will be kept or stored longer than needed.  However, in case of a script or
server fault or if the process is getting killed by the user, uploaded images
and results might not be deleted as intended and therefore stored in the
temporary folder of the corresponding server for a certain amount of time.

```{r,fig=FALSE,echo=TRUE}
library("colorspace")
print( rnorm(3) )
```