### 2019-01-10 pkg/colorspace/index.Rmd
employ ag_Synset (rather than Heat) on index page to correspond to logo/favicon
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@412 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

## A Toolbox for Manipulating and Assessing Colors and Palettes

```{r hcl, echo = FALSE, fig.show = "hide", fig.width = 4.5, fig.height = 4.5, fig.align = "center", dev = "png", dpi = 48}
library("colorspace")
hclplot(qualitative_hcl(9, c = 50, l = 70))
```

```{r palettes, echo = FALSE, fig.show = "hide", fig.width = 4.5, fig.height = 4.5, fig.align = "center", dev = "png", dpi = 48}
hcl_palettes(palette = c(
  "Pastel 1", "Dark 3", "Harmonic",
  "Blues 3", "Greens 3", "Reds 3",
  "ag_Sunset", "Purple-Blue", "Viridis",
  "Blue-Red", "Green-Brown", "Tropic"
), plot = TRUE, n = 7)
```

```{r ggplot2, echo = FALSE, fig.show = "hide", fig.width = 4.5, fig.height = 4.5, fig.align = "center", dev = "png", dpi = 48}
library("ggplot2")
theme_set(theme_minimal())
ggplot(iris, aes(x = Sepal.Length, fill = Species)) + geom_density(alpha = 0.6) + scale_fill_discrete_qualitative()
```

```{r inferno, echo = FALSE, fig.show = "hide", fig.width = 4.5, fig.height = 3.7, fig.align = "center", dev = "png", dpi = 48}
par(mar = rep(0, 4))
demoplot(sequential_hcl(99, "Inferno", rev = TRUE), type = "heatmap")
```
```{r deutan, echo = FALSE, fig.show = "hide", fig.width = 4.5, fig.height = 3.7, fig.align = "center", dev = "png", dpi = 48}
par(mar = rep(0, 4))
demoplot(deutan(sequential_hcl(99, "Inferno", rev = TRUE)), type = "heatmap")
```

```{r desaturate, echo = FALSE, fig.show = "hide", fig.width = 4.5, fig.height = 3.7, fig.align = "center", dev = "png", dpi = 48}
par(mar = rep(0, 4))
demoplot(desaturate(sequential_hcl(99, "Inferno", rev = TRUE)), type = "heatmap")
```

| Color spaces                               | HCL-based palettes                          | ggplot2 scales                                |
|:------------------------------------------:|:-------------------------------------------:|:---------------------------------------------:|
| ![](index_files/figure-html/hcl-1.png)     | ![](index_files/figure-html/palettes-1.png) | ![](index_files/figure-html/ggplot2-1.png)    |

| Interactive color apps                         |
|:----------------------------------------------:|
| ![](articles/screenshots/choose_palette_1.png) |

| Demos                                      | Color vision deficiency                     | Utilities                                     |
|:------------------------------------------:|:-------------------------------------------:|:---------------------------------------------:|
| ![](index_files/figure-html/inferno-1.png) | ![](index_files/figure-html/deutan-1.png)   | ![](index_files/figure-html/desaturate-1.png) |