### 2019-02-18 pkg/colorspace/paper/Toolbox/slides.Rnw
presentation slides (~30min) for 'toolbox' webpage/paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@447 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
%\usetheme[nototalframenumber, noslidenumber, license, url]{uibk}
\usetheme[license, url]{uibk}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      ]%
}
\headerimage{1}

\title{A Toolbox for Manipulating and Assessing Color Palettes for Statistical Graphics}
\subtitle{\phantom{.}\\Achim Zeileis, Jason C. Fisher, Kurt Hornik, Ross Ihaka, Claire D. McWhite, Paul Murrell, Reto Stauffer, Claus O. Wilke}
\author{\vspace*{-6.83cm}\hspace*{12.5cm}\includegraphics[width=1.6cm]{palmfooter.png}\\[5.5cm]}
\URL{http://colorspace.R-Forge.R-project.org/}

%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

<<preliminaries>>=
library("colorspace")
library("ggplot2")
theme_set(theme_minimal())
@

\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

\begin{document}

\section{colorspace}

\subsection{Color spaces}

\begin{frame}
\frametitle{Color spaces}

\textbf{Origin of the package:} Convert colors between various three-dimensional representations of color.

\medskip

\textbf{In particular:} From the perceptually-based HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB, and corresponding hex codes) space.

\setkeys{Gin}{width=0.8\textwidth}
\begin{center}
<<color-spaces, fig = TRUE, width = 9.5, height = 4.5>>=
par(mar = rep(0, 4))

d <- data.frame(
  x = c(1, 1, 2, 2, 3, 4, 5, 5, 5, 6),
  y = c(1, 3, 1, 3, 2, 2, 1, 3, 2, 2),
  lab = c("polarLAB", "polarLUV\n(= HCL)", "LAB", "LUV", "XYZ", "RGB", "HLS", "HSV", "sRGB", "hex"),
  stringsAsFactors = FALSE
)

plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
  xlim = c(0.7, 6.3), ylim = c(0.4, 3.6))
lines(y ~ x, data = d[-c(1, 3, 7, 8),])
lines(y ~ x, data = d[c(1, 3, 5),])
lines(y ~ x, data = d[c(6, 7, 9),])
lines(y ~ x, data = d[c(6, 8, 9),])
points(y ~ x, data = d[c(1, 2, 7, 8), ], pch = 21, bg = "lightgray", cex = 12)
points(y ~ x, data = d[c(3, 4, 5, 6, 9), ], pch = 22, bg = "lightgray", cex = 12)
points(y ~ x, data = d[10, ], pch = 23, bg = "white", cex = 12)

text(d$x, d$y, d$lab)
text(3.5, 2, "white point\n= D65", cex = 0.7, pos = 3)
text(4.56, 2, "gamma\n(= 2.4)", cex = 0.7, pos = 3)
@
\end{center}

\end{frame}

\subsection{HCL vs. RGB}

\begin{frame}
\frametitle{HCL vs.~RGB}

\begin{minipage}{0.47\textwidth}
\textbf{HCL:} Polar coordinates in CIELUV. Captures perceptual dimensions of the human visual system very well.

\bigskip

\setkeys{Gin}{width=\textwidth}
<<hcl-properties, fig = TRUE, width = 4, height = 2.2>>=
swatchplot(
  "Hue\n(Type of color)"       = sequential_hcl(5, h = c(0, 300), c = c(60, 60), l = 65),
  "Chroma\n(Colorfulness)"    = sequential_hcl(5, h = 0, c = c(100, 0), l = 65, rev = TRUE, power = 1),
  "Luminance\n(Brightness)" = sequential_hcl(5, h = 260, c = c(25, 25), l = c(25, 90), rev = TRUE, power = 1),
  nrow = 3, line = 5.2, off = 0
)
@
\end{minipage}
\pause
\hspace*{0.6cm}
\begin{minipage}{0.47\textwidth}
\textbf{RGB:} Motivated by how computers/TVs used to generate and still represent color.

\bigskip

\setkeys{Gin}{width=0.9\textwidth}
<<rgb-properties, fig = TRUE, width = 3.6, height = 2.2>>=
swatchplot(
  "Red"       = rgb(0:4/4, 0, 0),
  "Green"     = rgb(0, 0:4/4, 0),
  "Blue"      = rgb(0, 0, 0:4/4),
  nrow = 3, off = 0
)
@
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{HCL vs.~RGB: The End of the Rainbow}

\setkeys{Gin}{width=0.44\textwidth}
<<hcl-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow_hcl(99, c = 60, l = 75), rgb = TRUE)
@
\hspace*{0.8cm}
\pause
<<rgb-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow(95), rgb = TRUE)
@

\end{frame}

\subsection{Color palettes}

\begin{frame}
\frametitle{Color palettes: Somewhere over the Rainbow}

\setkeys{Gin}{width=\textwidth}
<<hcl-palettes-principles, fig = TRUE, width = 12, height = 1.6>>=
swatchplot(
  "Qualitative (Set 2)"     = rbind("Color" = qualitative_hcl(5, "Set 2"),     "Desaturated" = desaturate(qualitative_hcl(5, "Set 2"))),
  "Sequential (Blues 3)"    = rbind("Color" = sequential_hcl(7, "Blues 3"),    "Desaturated" = desaturate(sequential_hcl(7, "Blues 3"))),
  "Diverging (Green-Brown)" = rbind("Color" = diverging_hcl(7, "Green-Brown"), "Desaturated" = desaturate(diverging_hcl(7, "Green-Brown"))),
  nrow = 3, line = 7)
@

\bigskip
\pause

\textbf{Qualitative:} For categorical information, i.e., where no particular ordering of categories is available.
Function: \code{qualitative\_hcl()}.

\bigskip
\pause

\textbf{Sequential:} For ordered/numeric information, i.e., where colors go from high to low (or vice versa).
Function: \code{sequential\_hcl()}.

\bigskip
\pause

\textbf{Diverging:} For ordered/numeric information around a central neutral value, i.e.,
where colors diverge from neutral to two extremes. Function: \code{diverging\_hcl()}.

\end{frame}

\begin{frame}[fragile]
\frametitle{Color palettes: Somewhere over the Rainbow}

\only<1>{
\textbf{Sequential:} Luminance contrast is crucial (dark to light or vice versa).
}
\only<2>{
\textbf{Blues 2:} Single hue. Decreasing chroma with increasing luminance.
}
\only<3>{
\textbf{Blues 3:} Single hue. Triangular chroma to achieve higher luminance contrast.
}
\only<4>{
\textbf{Blues:} Multi hue. Triangular chroma. High luminance contrast.
}

\vspace*{-0.4cm}

\begin{minipage}[t]{0.49\textwidth}
\vspace*{0.46cm}

\setkeys{Gin}{width=\textwidth}
<<blues, fig = TRUE, width = 4, height = 2.2>>=
swatchplot(
  "Blues 2" = sequential_hcl(7, "Blues 2"),
  "Blues 3" = sequential_hcl(7, "Blues 3"),
  "Blues"   = sequential_hcl(7, "Blues")
)
@
\end{minipage}
\only<2>{%
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues2-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(sequential_hcl(7, "Blues 2"))
@
\end{minipage}%
}%
\only<3>{%
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues3-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(sequential_hcl(7, "Blues 3"))
@
\end{minipage}
}%
\only<4>{%
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(sequential_hcl(7, "Blues"))
@
\end{minipage}
}

\end{frame}

\begin{frame}[fragile]
\frametitle{Color palettes: Somewhere over the Rainbow}

\textbf{Diverging:} Combine two sequential palettes with balanced chroma/luminance.

\vspace*{-0.4cm}

\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<red-blue-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(diverging_hcl(9, "Blue-Red", h2 = 5, rev = TRUE))
@
\end{minipage}
\pause
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<green-brown-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(diverging_hcl(9, "Green-Brown"))
@
\end{minipage}

\end{frame}

\begin{frame}[fragile]
\frametitle{Color palettes: Somewhere over the Rainbow}

\setkeys{Gin}{width=\textwidth}
<<hcl-palettes, fig = TRUE, width = 12, height = 6, echo = TRUE>>=
hcl_palettes(plot = TRUE)
@

\end{frame}

\subsection{Statistical graphics}

\begin{frame}[fragile]
\frametitle{Statistical graphics}

\textbf{Base:}

\begin{itemize}
  \item HCL palette functions return hex color vector.
  \item Typically passed to \code{col = } argument of base plotting functions.
\end{itemize}

\bigskip

\textbf{ggplot2:}

\begin{itemize}
  \item Scales of type \code{scale_<aesthetic>_<datatype>_<colorscale>()}.
  \item \code{<aesthetic>} is \code{fill} or \code{color}/\code{colour}.
  \item \code{<datatype>} is \code{discrete} or \code{continuous}.
  \item \code{<colorscale>} is \code{qualitative}, \code{sequential}, \code{diverging}, or \code{divergingx}.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Statistical graphics: Base}

<<base-graphics, eval = FALSE, echo = TRUE>>=
q4 <- qualitative_hcl(4, palette = "Dark 3")
plot(log(EuStockMarkets), plot.type = "single", col = q4, lwd = 2)
legend("topleft", colnames(EuStockMarkets), col = q4, lwd = 3, bty = "n")
@

\vspace*{-0.6cm}

\setkeys{Gin}{width=0.8\textwidth}
<<base-graphics-plot, fig = TRUE, width = 10, height = 6>>=
<<base-graphics>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Statistical graphics: Base}

<<base-graphics2, eval = FALSE, echo = TRUE>>=
ttnc <- margin.table(Titanic, c(1, 4))[, 2:1]
spineplot(ttnc, col = sequential_hcl(2, palette = "Purples 3"))
@

\vspace*{-0.6cm}

\setkeys{Gin}{width=0.8\textwidth}
<<base-graphics2-plot, fig = TRUE, width = 10, height = 6>>=
<<base-graphics2>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Statistical graphics: ggplot2}

\setkeys{Gin}{width=0.8\textwidth}
<<ggplot2-graphics, fig = TRUE, width = 10, height = 5.5, echo = TRUE>>=
library("ggplot2")
ggplot(iris, aes(x = Sepal.Length, fill = Species)) + geom_density(alpha = 0.6) +
  scale_fill_discrete_qualitative(palette = "Dark 3")
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Statistical graphics: ggplot2}

\setkeys{Gin}{width=0.8\textwidth}
<<ggplot2-graphics2, fig = TRUE, width = 10, height = 5.5, echo = TRUE>>=
dsamp <- diamonds[1 + 1:1000 * 50, ]
ggplot(dsamp, aes(carat, price, color = cut)) + geom_point() +
  scale_color_discrete_sequential(palette = "Purples 3", nmax = 6, order = 2:6)
@

\end{frame}

\subsection{Visualization and assessment}

\begin{frame}[fragile]
\frametitle{Visualization and assessment}

\textbf{Visualizations:} Based on vector of colors.

\begin{itemize}
  \item \code{swatchplot()}: Color swatches.
  \item \code{specplot()}: Spectrum of HCL and/or RGB trajectories.
  \item \code{hclplot()}: Trajectories in 2-dimensional HCL space projections.
  \item \code{demoplot()}: Illustrations of typical (and simplified) statistical graphics.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Visualization and assessment: \texttt{hclplot()}}

<<hclplot, eval = FALSE, echo = TRUE>>=
hclplot(qualitative_hcl(7, palette = "Set 2"))
hclplot( sequential_hcl(7, palette = "Blues 3"))
hclplot(  diverging_hcl(7, palette = "Blue-Red"))
@

\medskip

\setkeys{Gin}{width=\textwidth}
<<hclplot-plot, fig = TRUE, width = 15, height = 5, png = TRUE, pdf = FALSE>>=
par(mfrow = c(1, 3))
<<hclplot>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Visualization and assessment: \texttt{demoplot()}}

<<demoplot-code, echo = TRUE, eval = FALSE>>=
cl <- sequential_hcl(5, palette = "Heat")
demoplot(cl, type = "...")
@

\medskip

\setkeys{Gin}{width=\textwidth}
<<demoplot, fig = TRUE, width = 15, height = 6>>=
par(mfrow = c(2, 5))
cl <- sequential_hcl(5, "Heat")
for (i in c("map", "heatmap", "scatter", "spine", "bar", "pie", "perspective", "mosaic", "lines")) {
  demoplot(cl, type = i)
  mtext(side = 3, text = i, line = 1.5)
}
@

\end{frame}

\subsection{Color vision deficiency}

\begin{frame}[fragile]
\frametitle{Color vision deficiency}

\textbf{Emulate:} Color vision deficies.

\begin{itemize}
  \item \code{deutan()}: Deuteranopia (green deficient).
  \item \code{protan()}: Protanopia (red deficient).
  \item \code{tritan()}: Tritanopia (blue deficient).
\end{itemize}

\bigskip

\textbf{Example:} Maunga Whau \code{volcano} data.

\vspace*{-2.2cm}

\setkeys{Gin}{width=0.6\textwidth}
<<volcano-persp, fig = TRUE, width = 5.5, height = 5.5>>=
z <- 2 * volcano
x <- 10 * (1:nrow(z))
y <- 10 * (1:ncol(z))
par(mar = rep(0, 4))
persp(x, y, z, theta = 10, phi = 30, col = "lightgray", scale = FALSE,
  ltheta = 100, shade = 0.5, border = NA, box = FALSE)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Color vision deficiency}

\setkeys{Gin}{width=\textwidth}
<<volcano-cvd, fig = TRUE, width = 15, height = 7>>=
par(mfrow = c(2, 3), mar = rep(0.5, 4), oma = c(1, 20, 3, 1), las = 1)
rain <- rainbow(11, end = 2/3)
blyl <- sequential_hcl(11, "Blue-Yellow")
demoplot(rain, "heatmap")
mtext("rainbow(11, end = 2/3)", side = 2, line = 1)
mtext("Original", line = 1)
demoplot(desaturate(rain), "heatmap")
mtext("Desaturated", line = 1)
demoplot(deutan(rain), "heatmap")
mtext("Deuteranope", line = 1)
demoplot(blyl, "heatmap")
mtext("sequential_hcl(11, \"Blue-Yellow\")", side = 2, line = 1)
demoplot(desaturate(blyl), "heatmap")
demoplot(deutan(blyl), "heatmap")
@

\end{frame}

\subsection{Approximations of other palettes}

\begin{frame}
\frametitle{Approximations of other palettes}

\vspace*{-0.1cm}

\setkeys{Gin}{width=0.49\textwidth}
<<ylgnbu-spectrum, fig = TRUE, width = 5.3, height = 5.6>>=
specplot(
  rev(RColorBrewer::brewer.pal(9, "YlGnBu")),
  sequential_hcl(9, "YlGnBu"),
  main = "ColorBrewer.org: YlGnBu"
)
@
\pause
\setkeys{Gin}{width=0.49\textwidth}
<<viridis-spectrum, fig = TRUE, width = 5.3, height = 5.6>>=
specplot(
  viridis::viridis(9),
  sequential_hcl(9, "Viridis"),
  main = "Viridis")
@

\end{frame}

\subsection{Color apps}

\begin{frame}[fragile]
\frametitle{Color apps}

\textbf{Facilitate exploration:} Graphical user interfaces as shiny apps.

\begin{itemize}
  \item \emph{Palette constructor:} \code{choose\_palette()} or \code{hclwizard()} (also in tcltk).
  \item \emph{Color picker:} \code{choose\_color()} or \code{hcl\_color\_picker()}.
  \item \emph{Color vision deficiency emulator:} \code{cvd\_emulator()}.
\end{itemize}

\bigskip

\textbf{Online versions:} \url{http://hclwizard.org/}

\end{frame}

\begin{frame}
\frametitle{Color apps: \texttt{choose\_palette()} / \texttt{hclwizard()}}

\includegraphics[width=0.9\textwidth]{../../vignettes/screenshots/choose_palette_1.png}

\end{frame}

\begin{frame}
\frametitle{Color apps: \texttt{choose\_color()} / \texttt{hcl\_color\_picker()}}

\includegraphics[width=0.65\textwidth]{../../vignettes/screenshots/choose_color.png}

\end{frame}

\begin{frame}
\frametitle{Color apps: \texttt{cvd\_emulator()}}

\includegraphics[width=0.34\textwidth]{../../vignettes/screenshots/cvd_emulator_1.png}
\includegraphics[width=0.34\textwidth]{../../vignettes/screenshots/cvd_emulator_2.png}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\small

Ihaka R, Murrell P, Hornik K, Fisher JC, Stauffer R, Wilke CO, McWhite CD, Zeileis A (2019).
  \emph{colorspace: A Toolbox for Manipulating and Assessing Colors and Palettes.}
  \url{http://colorspace.R-Forge.R-project.org/}

\bigskip

Zeileis A, Hornik K, Murrell P (2009).
 \dquote{Escaping RGBland: Selecting Colors for Statistical Graphics.}
 \textit{Computational Statistics \& Data Analysis}, \textbf{53}, 3259--3270.
 \doi{10.1016/j.csda.2008.11.033}.

\bigskip

Stauffer R, Mayr GJ, Dabernig M, Zeileis A (2015).
  \dquote{Somewhere over the Rainbow: How to Make Effective Use of Colors in Meteorological Visualizations.}
  \emph{Bulletin of the American Meteorological Society}, \textbf{96}(2), 203--216.
  \doi{10.1175/BAMS-D-13-00155.1}

\end{frame}

\end{document}