### 2020-09-22 pkg/colorspace/paper/ZAMG2020/slides.Rnw
Adding slides for ZAMG internal presentation, September 2020.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@555 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
%\usetheme[nototalframenumber, noslidenumber, license, url]{uibk}
\usetheme[nototalframenumber, license, url]{uibk}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      ]%
}
\headerimage{1}

\title{Robust Color Maps That Work for Most Audiences (Including the U.S. President)}
\subtitle{Reto Stauffer, Achim Zeileis}
\author{\vspace*{-6.1cm}\hspace*{12.5cm}\includegraphics[width=1.6cm]{palmfooter.png}\\[5.5cm]\vspace{-1.5em}} %% -6.83 with "for Statistical Graphics"
\URL{http://hclwizard.org/}%

%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

<<preliminaries>>=
library("colorspace")
@

\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

\begin{document}

\section{colorspace: A Toolbox for Manipulating and Assessing Color Palettes} %

\subsection{Motivation}%

% ---------------------------------------------------------------------
\begin{frame}
\frametitle{Motivation}

\vspace*{2em}

\begin{minipage}{0.46\textwidth}
    \includegraphics[width=\textwidth]{images/sharpie-small.png}
\end{minipage}
\pause
\hfill
\begin{minipage}{0.51\textwidth}
    \includegraphics[width=\textwidth]{images/tweet.png}
\end{minipage}

    \vspace*{0.5em}
    \begin{footnotesize}
    Projected track and wind speed forecast of hurricane Dorian.
    Screenshot of a video released by the White House (Sep. 4, 2019),
    tweet by the U.S. president (Sep. 5, 2019).
    \end{footnotesize}

\end{frame}

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Motivation}
%
%\vspace*{-0.3cm}
%
%\begin{minipage}{0.54\textwidth}
%    \only<1>{\includegraphics[width=\textwidth]{images/dorian-rainbow.png}}%
%    \only<2-3>{\includegraphics[width=\textwidth]{images/dorian-rainbow-arrow.png}}%
%    \only<4>{\includegraphics[width=\textwidth]{images/dorian-rainbow-deutan.png}}%
%    \only<5>{\includegraphics[width=\textwidth]{images/dorian-rainbow-gray.png}}%
%    \only<6>{\includegraphics[width=\textwidth]{images/dorian-hclrainbow.png}}%
%    \only<7>{\includegraphics[width=\textwidth]{images/dorian-hclrainbow-gray.png}}
%    \only<8>{\includegraphics[width=\textwidth]{images/dorian-hclrainbow-deutan.png}}
%\end{minipage}
%\hspace*{0.6cm}
%\begin{minipage}{0.4\textwidth}
%
%    \textbf{Map:} Probability of exceeding wind speeds of 63\,km\,h$^{-1}$,
%    Aug~30 -- Sep~04, 2019.
%
%    \medskip
%    \textbf{Source:} NOAA (\href{https://noaa.gov}{noaa.gov})
%
%    \medskip
%    \visible<3->{\textbf{Problems:} Flashy.}
%    \visible<4->{Color vision deficiency.}
%    \visible<5->{Grayscale.}
%    
%    \medskip
%    \visible<6->{\textbf{Alternative:} HCL-based sequential palette.}
%
%\end{minipage}
%
%\end{frame}

% ---------------------------------------------------------------------
\begin{frame}
    \frametitle{\only<1-4>{Problem}\only<5->{Solution}}

\vspace*{-0.3cm}

    \begin{center}
        \only<1>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow.png}}%
        \only<2>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow-arrow.png}}%
        \only<3>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow-deutan.png}}%
        \only<4>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow-gray.png}}%
        \only<5>{\includegraphics[width=.55\textwidth]{images/dorian-hclrainbow.png}}%
        \only<6>{\includegraphics[width=.55\textwidth]{images/dorian-hclrainbow-gray.png}}
        \only<7>{\includegraphics[width=.55\textwidth]{images/dorian-hclrainbow-deutan.png}}
    \end{center}

    \vspace*{-0.5em}
    \begin{footnotesize} 
    Probability of exceeding wind speeds of 39\,mph (63\,km\,h$^{-1}$),
    Aug~30 -- Sep~04, 2019 (\href{https://noaa.gov}{noaa.gov}).
    \end{footnotesize}

\end{frame}

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Motivation}
%
%\textbf{Need tools for:}
%\begin{itemize}
%  \item Construction of palettes with better perceptual properties.
%  \item Assessment of color palettes.
%  \item Manipulation of colors.
%\end{itemize}
%
%\end{frame}

\subsection{Color spaces}

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Color spaces}
%
%\textbf{Origin of the package:} Convert colors between various three-dimensional representations of color.
%
%\medskip
%
%\textbf{In particular:} From the perceptually-based HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB, and corresponding hex codes) space.
%
%\setkeys{Gin}{width=0.8\textwidth}
%\begin{center}
%<<color-spaces, fig = TRUE, width = 9.5, height = 4.5>>=
%par(mar = rep(0, 4))
%
%d <- data.frame(
%  x = c(1, 1, 2, 2, 3, 4, 5, 5, 5, 6),
%  y = c(1, 3, 1, 3, 2, 2, 1, 3, 2, 2),
%  lab = c("polarLAB", "polarLUV\n(= HCL)", "LAB", "LUV", "XYZ", "RGB", "HLS", "HSV", "sRGB", "hex"),
%  stringsAsFactors = FALSE
%)
%
%plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
%  xlim = c(0.7, 6.3), ylim = c(0.4, 3.6))
%lines(y ~ x, data = d[-c(1, 3, 7, 8),])
%lines(y ~ x, data = d[c(1, 3, 5),])
%lines(y ~ x, data = d[c(6, 7, 9),])
%lines(y ~ x, data = d[c(6, 8, 9),])
%points(y ~ x, data = d[c(1, 2, 7, 8), ], pch = 21, cex = 12,
%       bg = c("lightgray", "#F5CF8B", "lightgray", "lightgray"))
%points(y ~ x, data = d[c(3, 4, 5, 6, 9), ], pch = 22, cex = 12,
%       bg = ifelse(1:5 == 5, "#F5CF8B", "lightgray"))
%points(y ~ x, data = d[10, ], pch = 23, bg = "white", cex = 12)
%
%text(d$x, d$y, d$lab)
%text(3.5, 2, "white point\n= D65", cex = 0.7, pos = 3)
%text(4.56, 2, "gamma\n(= 2.4)", cex = 0.7, pos = 3)
%@
%\end{center}
%
%\end{frame}

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Color spaces}
%
%\textbf{Origin of the package:} Convert colors between various three-dimensional representations of color.
%
%\medskip
%
%\textbf{In particular:} From the perceptually-based HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB, and corresponding hex codes) space.
%
%\setkeys{Gin}{width=0.8\textwidth}
%\begin{center}
%<<color-spaces, fig = TRUE, width = 9.5, height = 4.5>>=
%par(mar = rep(0, 4))
%
%d <- data.frame(
%  x = c(1, 1, 2, 2, 3, 4, 5, 5, 5, 6),
%  y = c(1, 3, 1, 3, 2, 2, 1, 3, 2, 2),
%  lab = c("polarLAB", "polarLUV\n(= HCL)", "LAB", "LUV", "XYZ", "RGB", "HLS", "HSV", "sRGB", "hex"),
%  stringsAsFactors = FALSE
%)
%
%plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
%  xlim = c(0.7, 6.3), ylim = c(0.4, 3.6))
%lines(y ~ x, data = d[-c(1, 3, 7, 8),])
%lines(y ~ x, data = d[c(1, 3, 5),])
%lines(y ~ x, data = d[c(6, 7, 9),])
%lines(y ~ x, data = d[c(6, 8, 9),])
%points(y ~ x, data = d[c(1, 2, 7, 8), ], pch = 21, bg = "lightgray", cex = 12)
%points(y ~ x, data = d[c(3, 4, 5, 6, 9), ], pch = 22, bg = "lightgray", cex = 12)
%points(y ~ x, data = d[10, ], pch = 23, bg = "white", cex = 12)
%
%text(d$x, d$y, d$lab)
%text(3.5, 2, "white point\n= D65", cex = 0.7, pos = 3)
%text(4.56, 2, "gamma\n(= 2.4)", cex = 0.7, pos = 3)
%@
%\end{center}
%
%\end{frame}

\subsection{HCL vs. RGB}

% ---------------------------------------------------------------------
\begin{frame}
\frametitle{HCL vs.~RGB}

\begin{minipage}{0.47\textwidth}
\textbf{HCL:} Polar coordinates in CIELUV. Captures perceptual dimensions of the human visual system very well.

\bigskip

\setkeys{Gin}{width=\textwidth}
<<hcl-properties, fig = TRUE, width = 4, height = 2.2>>=
swatchplot(
  "Hue\n(Type of color)"       = sequential_hcl(5, h = c(0, 300), c = c(60, 60), l = 65),
  "Chroma\n(Colorfulness)"    = sequential_hcl(5, h = 0, c = c(100, 0), l = 65, rev = TRUE, power = 1),
  "Luminance\n(Brightness)" = sequential_hcl(5, h = 260, c = c(25, 25), l = c(25, 90), rev = TRUE, power = 1),
  nrow = 3, line = 5.2, off = 0
)
@
\end{minipage}
\pause
\hspace*{0.6cm}
\begin{minipage}{0.47\textwidth}
\textbf{RGB:} Motivated by how computers/TVs used to generate and still represent color.

\bigskip

\setkeys{Gin}{width=0.9\textwidth}
<<rgb-properties, fig = TRUE, width = 3.6, height = 2.2>>=
swatchplot(
  "Red"       = rgb(0:4/4, 0, 0),
  "Green"     = rgb(0, 0:4/4, 0),
  "Blue"      = rgb(0, 0, 0:4/4),
  nrow = 3, off = 0
)
@
\end{minipage}

\end{frame}

% ---------------------------------------------------------------------
\begin{frame}
\frametitle{HCL vs.~RGB}

\setkeys{Gin}{width=1.0\textwidth}

<<specplot-init, include = FALSE>>=
orig <- c("#008a02", "#00cb00", "#82fd00", "#f5f900", "#f7d600",
          "#d28400", "#fb7502", "#a80003", "#740001", "#8c0084")
hcl  <- c("#b0ceb9", "#9fca91", "#a5bc62", "#b3a924", "#c09000",
          "#c87400", "#c85435", "#b9375c", "#a4176e", "#890075")
@

    \begin{minipage}{0.48\textwidth}
<<specplot-orig, echo = FALSE, fig = TRUE, fig.width = 5, fig.height = 4>>=
specplot(orig, rgb = FALSE, main = "Problem")
@
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
<<specplot-hcl, echo = FALSE, fig = TRUE, fig.width = 5, fig.height = 4>>=
specplot(hcl, rgb = FALSE, main = "Solution")
@
    \end{minipage}

\end{frame}

\begin{frame}
\frametitle{HCL vs.~RGB: The End of the Rainbow}

\setkeys{Gin}{width=0.44\textwidth}
<<rgb-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow(95), rgb = TRUE)
@
\hspace*{0.8cm}
\pause
<<hcl-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow_hcl(99, c = 60, l = 75), rgb = TRUE)
@

\end{frame}

\subsection{Color palettes}

\begin{frame}
\frametitle{Color palettes: Somewhere over the Rainbow}

\setkeys{Gin}{width=\textwidth}
<<hcl-palettes-principles, fig = TRUE, width = 12, height = 1.6>>=
swatchplot(
  "Qualitative (Set 2)"     = rbind("Color" = qualitative_hcl(5, "Set 2"),     "Desaturated" = desaturate(qualitative_hcl(5, "Set 2"))),
  "Sequential (Blues 3)"    = rbind("Color" = sequential_hcl(7, "Blues 3"),    "Desaturated" = desaturate(sequential_hcl(7, "Blues 3"))),
  "Diverging (Green-Brown)" = rbind("Color" = diverging_hcl(7, "Green-Brown"), "Desaturated" = desaturate(diverging_hcl(7, "Green-Brown"))),
  nrow = 3, line = 7)
@

\bigskip
\pause

\textbf{Qualitative:} For categorical information, i.e., where no particular ordering of categories is available.
Function: \code{qualitative\_hcl()}.

\pause
\bigskip

\textbf{Sequential:} For ordered/numeric information, i.e., where colors go from high to low (or vice versa).
Function: \code{sequential\_hcl()}.

\pause
\bigskip

\textbf{Diverging:} For ordered/numeric information around a central neutral value, i.e.,
where colors diverge from neutral to two extremes.
Function: \code{diverging\_hcl()}.

\end{frame}

\begin{frame}[fragile]
\frametitle{Color palettes: Somewhere over the Rainbow}

\begin{minipage}[t]{0.49\textwidth}
\textbf{Blues 2:} Single hue. Decreasing chroma with increasing luminance.

\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues2-spectrum, fig = TRUE, width = 5.2, height = 4.8>>=
specplot(sequential_hcl(7, "Blues 2"))
@
\end{minipage}%
\hfill
\begin{minipage}[t]{0.49\textwidth}
\textbf{Blues 3:} Single hue. Triangular chroma to achieve higher luminance contrast.

\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues3-spectrum, fig = TRUE, width = 5.2, height = 4.8>>=
specplot(sequential_hcl(7, "Blues 3"))
@
\end{minipage}

\end{frame}

\begin{frame}[fragile]
\frametitle{Color palettes: Somewhere over the Rainbow}

\textbf{Diverging:} Combine two sequential palettes with balanced chroma/luminance.

\vspace*{-0.4cm}

\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<red-blue-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(diverging_hcl(9, "Blue-Red", h2 = 5, rev = TRUE))
@
\end{minipage}
\pause
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<green-brown-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(diverging_hcl(9, "Green-Brown"))
@
\end{minipage}

\end{frame}

\begin{frame}[fragile]
\frametitle{Color palettes: Somewhere over the Rainbow}

\setkeys{Gin}{width=\textwidth}
<<hcl-palettes, fig = TRUE, width = 12, height = 6, echo = TRUE>>=
hcl_palettes(plot = TRUE)
@

\end{frame}

\subsection{Visualization and assessment}

\begin{frame}[fragile]
\frametitle{Visualization and assessment}

\textbf{Visualizations:} Based on vector of colors.

\begin{itemize}
  \item \code{swatchplot()}: Color swatches.
  \item \code{specplot()}: Spectrum of HCL and/or RGB trajectories.
  \item \code{hclplot()}: Trajectories in 2-dimensional HCL space projections.
  \item \code{demoplot()}: Illustrations of typical (and simplified) statistical graphics.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Visualization and assessment: \texttt{hclplot()}}

<<hclplot, eval = FALSE, echo = TRUE>>=
hclplot(qualitative_hcl(7, palette = "Set 2"))
hclplot( sequential_hcl(7, palette = "Blues 3"))
hclplot(  diverging_hcl(7, palette = "Blue-Red"))
@

\medskip

\setkeys{Gin}{width=\textwidth}
<<hclplot-plot, fig = TRUE, width = 15, height = 5, png = TRUE, pdf = FALSE>>=
par(mfrow = c(1, 3))
<<hclplot>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Visualization and assessment: \texttt{demoplot()}}

<<demoplot-code, echo = TRUE, eval = FALSE>>=
cl <- sequential_hcl(5, palette = "Heat")
demoplot(cl, type = "...")
@

\medskip

\setkeys{Gin}{width=\textwidth}
<<demoplot, fig = TRUE, width = 15, height = 6>>=
par(mfrow = c(2, 5))
cl <- sequential_hcl(5, "Heat")
for (i in c("map", "heatmap", "scatter", "spine", "bar", "pie", "perspective", "mosaic", "lines")) {
  demoplot(cl, type = i)
  mtext(side = 3, text = i, line = 1.5)
}
@

\end{frame}

\subsection{Color vision deficiency}

\begin{frame}[fragile]
\frametitle{Color vision deficiency}

\textbf{Emulate:} Color vision deficiencies.

\begin{itemize}
  \item \code{deutan()}: Deuteranopia (green deficient).
  \item \code{protan()}: Protanopia (red deficient).
  \item \code{tritan()}: Tritanopia (blue deficient).
\end{itemize}

\bigskip

\textbf{Example:} Maunga Whau \code{volcano} data.

\vspace*{-2.2cm}

\setkeys{Gin}{width=0.6\textwidth}
<<volcano-persp, fig = TRUE, width = 5.5, height = 5.5>>=
z <- 2 * volcano
x <- 10 * (1:nrow(z))
y <- 10 * (1:ncol(z))
par(mar = rep(0, 4))
persp(x, y, z, theta = 10, phi = 30, col = "lightgray", scale = FALSE,
  ltheta = 100, shade = 0.5, border = NA, box = FALSE)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Color vision deficiency}

\setkeys{Gin}{width=\textwidth}
<<volcano-cvd, fig = TRUE, width = 15, height = 7>>=
par(mfrow = c(2, 3), mar = rep(0.5, 4), oma = c(1, 20, 3, 1), las = 1)
rain <- rainbow(11, end = 2/3)
blyl <- sequential_hcl(11, "Blue-Yellow")
demoplot(rain, "heatmap")
mtext("rainbow(11, end = 2/3)", side = 2, line = 1)
mtext("Original", line = 1)
demoplot(desaturate(rain), "heatmap")
mtext("Desaturated", line = 1)
demoplot(deutan(rain), "heatmap")
mtext("Deuteranope", line = 1)
demoplot(blyl, "heatmap")
mtext("sequential_hcl(11, \"Blue-Yellow\")", side = 2, line = 1)
demoplot(desaturate(blyl), "heatmap")
demoplot(deutan(blyl), "heatmap")
@

\end{frame}

\subsection{Approximations of other palettes}

\begin{frame}
\frametitle{Approximations of other palettes}

\vspace*{-0.1cm}

\setkeys{Gin}{width=0.49\textwidth}
<<ylgnbu-spectrum, fig = TRUE, width = 5.3, height = 5.6>>=
specplot(
  rev(RColorBrewer::brewer.pal(9, "YlGnBu")),
  sequential_hcl(9, "YlGnBu"),
  main = "ColorBrewer.org: YlGnBu"
)
@
\pause
\setkeys{Gin}{width=0.49\textwidth}
<<viridis-spectrum, fig = TRUE, width = 5.3, height = 5.6>>=
specplot(
  viridis::viridis(9),
  sequential_hcl(9, "Viridis"),
  main = "Viridis")
@

\end{frame}

\subsection{Base R}

\begin{frame}
\frametitle{Base R}

\textbf{In 3.6.0:} All prespecified palettes also via \code{grDevices::hcl.colors()}.

\vspace*{-1.8cm}

\setkeys{Gin}{width=\textwidth}
<<hcl-colors, fig = TRUE, width = 11, height = 8>>=
hcl.swatch <- function(palette = hcl.pals(), n = 5, nrow = 11,
  border = if (n < 15) "black" else NA)
{
    cols <- sapply(palette, hcl.colors, n = n)
    ncol <- ncol(cols)
    nswatch <- min(ncol, nrow)

    par(mar = rep(0.1, 4),
        mfrow = c(1, min(10, ncol %/% nrow + 1)),
        pin = c(1, 0.5 * nswatch),
        cex = 0.7)

    while (length(palette)) {
        subset <- 1:min(nrow, ncol(cols))
        plot.new()
        plot.window(c(0, n), c(0, nrow + 1))
        text(0, rev(subset) + 0.1, palette[subset], adj = c(0, 0))
        y <- rep(subset, each = n)
        rect(rep(0:(n-1), n), rev(y), rep(1:n, n), rev(y) - 0.5,
             col = cols[, subset], border = border)
        palette <- palette[-subset]
        cols <- cols[, -subset]
    }

    par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1), cex = 1)
}
hcl.swatch()
@

\end{frame}

\subsection{Color apps}

\begin{frame}[fragile]
\frametitle{Color apps}

\textbf{Facilitate exploration:} Graphical user interfaces as shiny apps.

\begin{itemize}
  \item \emph{Palette constructor:} \code{choose\_palette()} or \code{hclwizard()} (also in tcltk).
  \item \emph{Color picker:} \code{choose\_color()} or \code{hcl\_color\_picker()}.
  \item \emph{Color vision deficiency emulator:} \code{cvd\_emulator()}.
\end{itemize}

\bigskip

\textbf{Online versions:} \url{http://hclwizard.org/}

\end{frame}

\begin{frame}
\frametitle{Color apps: \texttt{choose\_palette()} / \texttt{hclwizard()}}

\includegraphics[width=0.9\textwidth]{../../vignettes/screenshots/choose_palette_1.png}

\end{frame}

\begin{frame}
\frametitle{Color apps: \texttt{choose\_color()} / \texttt{hcl\_color\_picker()}}

\includegraphics[width=0.65\textwidth]{../../vignettes/screenshots/choose_color.png}

\end{frame}

\begin{frame}
\frametitle{Color apps: \texttt{cvd\_emulator()}}

\includegraphics[width=0.34\textwidth]{../../vignettes/screenshots/cvd_emulator_1.png}
\includegraphics[width=0.34\textwidth]{../../vignettes/screenshots/cvd_emulator_2.png}

\end{frame}

\subsection{Recommendations}

\begin{frame}
\frametitle{Recommendations}

\textbf{Colors and palettes:}
\begin{itemize}
  \item Do not overestimate the effectiveness of color.
  \item Choose type of palette based on the data to be visualized.
  \item For areas use light colors (higher luminance, lower chroma).
  \item For points/lines darker colors are needed (lower luminance, higher chroma).
  \item For palettes with more colors stronger luminance contrasts are needed.
  \item Triangular chroma trajectories useful for distinguishing central colors.
\end{itemize}

\bigskip
\pause

\textbf{R packages:}
\begin{itemize}
  \item \pkg{colorspace} facilitates exploration, manipulation, and assessment.
  \item HCL approximations of palettes from \pkg{RColorBrewer}, \pkg{rcartocolor}, \pkg{scico}, \dots
  \item Prespecified palettes are also easily available in base R.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Do it yourself}

    \textbf{Software:} \emph{colorspace}.
    \begin{itemize}
      \item R (mature): \url{http://colorspace.R-Forge.R-project.org/}.
      \item Python (beta): \url{https://github.com/retostauffer/python-colorspace}.
      \item Web (interactive): \url{http://www.hclwizard.org/}.
    \end{itemize}

    \bigskip

    \textbf{Take-home messages:}
    \begin{itemize}
      \item Choose colors carefully.
      \item Make areas of interest stand out from background.
      \item Check robustness.
      \item Software helps you.
    \end{itemize}

\end{frame}

\subsection{References}

\begin{frame}\label{lastslide}
\frametitle{References}

\small

Zeileis A, Fisher JC, Hornik K, Ihaka R, McWhite CD, Murrell P, Stauffer R, Wilke CO (2019).
  \dquote{colorspace: A Toolbox for Manipulating and Assessing Colors and Palettes.}
  arXiv:1903.06490, arXiv.org E-Print Archive.
  \url{http://arxiv.org/abs/1903.06490}

\bigskip

Zeileis A, Hornik K, Murrell P (2009).
 \dquote{Escaping RGBland: Selecting Colors for Statistical Graphics.}
 \textit{Computational Statistics \& Data Analysis}, \textbf{53}, 3259--3270.
 \doi{10.1016/j.csda.2008.11.033}.

\bigskip

Stauffer R, Mayr GJ, Dabernig M, Zeileis A (2015).
  \dquote{Somewhere over the Rainbow: How to Make Effective Use of Colors in Meteorological Visualizations.}
  \emph{Bulletin of the American Meteorological Society}, \textbf{96}(2), 203--216.
  \doi{10.1175/BAMS-D-13-00155.1}

\end{frame}

\begin{frame}
	\frametitle{Thank you for your interest!}
	\begin{center}
		\includegraphics[width = .6\textwidth]{images/bobross.jpg}
	\end{center}
\end{frame}

\end{document}