### 2021-07-02 pkg/colorspace/paper/useR-2021/slides.Rnw
slides for useR-2021 keynote
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@601 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[license, url, foot]{uibk}
\renewcommand{\hrulefill}{}

\title{Choosing Color Palettes for Data Visualization}
\subtitle{Tools and Technologies for Supporting Algorithm Fairness and Inclusion}
\author{Achim Zeileis}
\URL{https://www.zeileis.org/}

%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

<<preliminaries>>=
library("colorspace")
eu <- window(EuStockMarkets, end = 1998)
@

\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

\begin{document}

\section{Robust Color Palettes for Data Visualization}

\subsection{Motivation}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Colors in data visualization:}
\begin{itemize}
  \item Ubiquitous.
  \item Not always easy to choose.
  \item But also perceived as fun.
\end{itemize}

\bigskip
\pause

\textbf{Potential problems:}
\begin{itemize}
  \item Power of color often overestimated.
  \item Color vision deficiencies affect $\sim 8$\% of male and $\sim 0.5$\% of female viewers.
  \item Other physical or technical limitations.
\end{itemize}

\bigskip
\pause

\textbf{Illustration:} Time series line plot using base graphics.

<<echo=TRUE,eval=FALSE>>=
p <- c(1:3, 5)
plot(EuStockMarkets, log = "y", plot.type = "single", col = p, ...)
legend("topleft", colnames(EuStockMarkets), col = p, ...)
@

\end{frame}

\begin{frame}
\frametitle{Motivation}

\begin{minipage}[t]{.73\textwidth}
\setkeys{Gin}{width=\textwidth}
\vspace*{-0.6cm}

<<EuStockMarkets0, eval=FALSE, echo=FALSE, results=hide>>=
par(mar = c(5, 4, 1, 1))
plot(eu, log = "y", ylab = "EU Stock Prices", plot.type = "single", col = p, xlim = c(1991.6, 1998.4), lwd = 1.5)
@
<<EuStockMarketsLegend1, eval=FALSE, echo=FALSE, results=hide>>=
legend("topleft", colnames(eu), col = p, lty = 1, lwd = 3, bty = "n")
@
<<EuStockMarketsLegend2, eval=FALSE, echo=FALSE, results=hide>>=
text(1998, tail(eu, 1), colnames(eu), pos = 4, col = p, xpd = TRUE)
@

\only<1-2>{%
<<EuStockMarkets1, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- palette.colors(8, "R3")[c(1:3, 5)]
<<EuStockMarkets0>>
<<EuStockMarketsLegend1>>
@
}%
\only<3>{%
<<EuStockMarkets2, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- protan(palette.colors(8, "R3")[c(1:3, 5)])
<<EuStockMarkets0>>
<<EuStockMarketsLegend1>>
@
}%
\only<4>{%
<<EuStockMarkets3, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- protan(palette.colors(8, "R3")[c(1:3, 5)])
<<EuStockMarkets0>>
<<EuStockMarketsLegend2>>
@
}%
\only<5>{%
<<EuStockMarkets4, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- palette.colors(8, "R3")[c(1:3, 5)]
<<EuStockMarkets0>>
<<EuStockMarketsLegend2>>
@
}%
\only<6>{%
<<EuStockMarkets5, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- palette.colors(8, "R4")[c(1:3, 5)]
<<EuStockMarkets0>>
<<EuStockMarketsLegend2>>
@
}%
\only<7>{%
<<EuStockMarkets6, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- protan(palette.colors(8, "R4")[c(1:3, 5)])
<<EuStockMarkets0>>
<<EuStockMarketsLegend2>>
@
}%
\only<8>{%
<<EuStockMarkets7, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- palette.colors(4, "Okabe-Ito")[c(1, 4, 2, 3)]
<<EuStockMarkets0>>
<<EuStockMarketsLegend2>>
@
}%
\only<9>{%
<<EuStockMarkets8, fig=TRUE, height=6, width=8.5, echo=FALSE, results=hide>>=
p <- protan(palette.colors(4, "Okabe-Ito")[c(1, 4, 2, 3)])
<<EuStockMarkets0>>
<<EuStockMarketsLegend2>>
@
}%

\end{minipage}
\hfill
\begin{minipage}[t]{0.26\textwidth}

\small

\textbf{Palette:}
\only<1-5>{R $\le$ 3 default}%
\only<6-7>{R $\ge$ 4 default}%
\only<8-9>{Okabe-Ito}

\medskip

\textbf{Emulation:}
\only<1-2,5-6,8>{None}%
\only<3-4,7,9>{Protanope}

\medskip

\textbf{Labeling:}
\only<1-3>{Legend}%
\only<4-9>{Direct}

\bigskip
\bigskip

\only<2->{\textbf{Comments:}}

\only<2-5>{
\medskip

Too flashy

\medskip

Cyan too light

}

\only<3-5>{

\medskip

Hard to distinguish for protanope viewers

}

\only<6-7>{
\medskip

Similar hues

\medskip

More balanced

brightness

\medskip

Avoid garish colors

}

\only<8-9>{
\medskip

Designed to be robust

against color vision

deficiencies

}

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Motivation}

%% png("drake.png", height = 570, width = 530, res = 100)
%% contrast_ratio(rev(palette.colors(8, "R3")), rev(palette.colors(8, "R3")), plot = TRUE, main = "old base palette", mar = c(1, 1, 3, 5), off = 0.4)
%% text(0, 7:0 + 0.73, palette.colors(8, "R3"), pos = 4)
%% 
%% png("drake.png", height = 570, width = 530, res = 100)
%% contrast_ratio(rev(palette.colors(8, "R4")), rev(palette.colors(8, "R4")), plot = TRUE, main = "new base palette", mar = c(1, 1, 3, 5), off = 0.4)
%% text(0, 7:0 + 0.73, palette.colors(8, "R4"), pos = 4)
%% dev.off()

\begin{minipage}[t]{0.51\textwidth}

\vspace*{-0.6cm}

\includegraphics[width=\textwidth]{images/palette-bling.png}

{\footnotesize
Source: Mara Averick via Twitter
}

\end{minipage}
\hspace*{0.3cm}
\pause
\begin{minipage}[t]{0.44\textwidth}

\textbf{Base R:} Neglected better color palettes for a long time.

\bigskip

\textbf{Earlier packages:} RColorBrewer, colorspace, ggplot2, viridis, rcartocolor, Polychrome, pals, paletteer, \dots

\bigskip

\textbf{Thus:} Many good palettes easily available.

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Motivation}

\includegraphics[width=1.0\textwidth]{images/grdevices-qualitative.png}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\setkeys{Gin}{width=0.6\textwidth}

<<swatchplot-r4, fig=TRUE, height=4, width=6.5, echo=TRUE, results=hide>>=
palette.colors(palette = "R4") |>
  colorspace::swatchplot(cvd = TRUE)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\setkeys{Gin}{width=0.6\textwidth}

<<swatchplot-okabeito, fig=TRUE, height=4, width=6.5, echo=TRUE, results=hide>>=
palette.colors(palette = "Okabe-Ito") |>
  colorspace::swatchplot(cvd = TRUE)
@

\end{frame}

\subsection{Color palette construction}

\begin{frame}
\frametitle{Color palette construction}

\begin{minipage}{0.47\textwidth}
\textbf{HCL:} Polar coordinates in CIELUV. Captures perceptual dimensions of the human visual system very well.

\bigskip

\setkeys{Gin}{width=\textwidth}
<<hcl-properties, fig = TRUE, width = 4, height = 2.2>>=
swatchplot(
  "Hue\n(Type of color)"       = sequential_hcl(5, h = c(0, 300), c = c(60, 60), l = 65),
  "Chroma\n(Colorfulness)"    = sequential_hcl(5, h = 0, c = c(100, 0), l = 65, rev = TRUE, power = 1),
  "Luminance\n(Brightness)" = sequential_hcl(5, h = 260, c = c(25, 25), l = c(25, 90), rev = TRUE, power = 1),
  nrow = 3, line = 5.2, off = 0
)
@
\end{minipage}
\pause
\hspace*{0.6cm}
\begin{minipage}{0.47\textwidth}
\textbf{RGB:} Motivated by how computers/TVs used to generate and still represent color.

\bigskip

\setkeys{Gin}{width=0.9\textwidth}
<<rgb-properties, fig = TRUE, width = 3.6, height = 2.2>>=
swatchplot(
  "Red"       = rgb(0:4/4, 0, 0),
  "Green"     = rgb(0, 0:4/4, 0),
  "Blue"      = rgb(0, 0, 0:4/4),
  nrow = 3, off = 0
)
@
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Color palette construction}

\setkeys{Gin}{width=\textwidth}
<<hcl-palettes-principles, fig = TRUE, width = 12, height = 1.6>>=
swatchplot(
  "Qualitative (Set 2)"     = rbind("Color" = qualitative_hcl(5, "Set 2"),     "Desaturated" = desaturate(qualitative_hcl(5, "Set 2"))),
  "Sequential (Blues 3)"    = rbind("Color" = sequential_hcl(7, "Blues 3"),    "Desaturated" = desaturate(sequential_hcl(7, "Blues 3"))),
  "Diverging (Green-Brown)" = rbind("Color" = diverging_hcl(7, "Green-Brown"), "Desaturated" = desaturate(diverging_hcl(7, "Green-Brown"))),
  nrow = 3, line = 7)
@

\bigskip
\pause

\textbf{Qualitative:} For categorical information with no particular ordering.

Luminance differences should be limited

\bigskip
\pause

\textbf{Sequential:} For ordered/numeric information from high to low (or vice versa).

\bigskip
\pause

\textbf{Diverging:} For ordered/numeric information diverging from a central neutral value to two extremes.

\end{frame}

\begin{frame}
\frametitle{Color palette construction}

\begin{minipage}[t]{0.47\textwidth}
\vspace*{-0.9cm}
\includegraphics[width=\textwidth]{images/grdevices-sequential.png}
\end{minipage}
\begin{minipage}[t]{0.39\textwidth}
\vspace*{-0.8cm}
\includegraphics[width=\textwidth]{images/grdevices-diverging.png}
\end{minipage}

\end{frame}

\subsection{Risk maps and communication to the public}

\begin{frame}
\frametitle{Risk maps and communication to the public}

\vspace*{-0.3cm}

\begin{center}
\only<1>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-orig.png}}%
\only<2>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-orig-deutan.png}}%
\only<3>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-orig-gray.png}}%
\only<4>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-orrd.png}}%
\only<5>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-orrd-gray.png}}
\only<6>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-orrd-deutan.png}}
\end{center}

\vspace*{-0.5em}
\begin{footnotesize} 
Probability of wind speeds $>$ 39\,mph (63\,km\,h$^{-1}$), 2019-08-30--2019-09-04.
Source: (\href{https://noaa.gov}{noaa.gov})
\end{footnotesize}

\end{frame}

\begin{frame}
\frametitle{Risk maps and communication to the public}

\vspace*{2em}

\begin{minipage}{0.46\textwidth}
\includegraphics[width=\textwidth]{../EGU2020/images/sharpie-small.png}

{\footnotesize
Source: White House (2019-09-04)}
\end{minipage}
\hfill
\begin{minipage}{0.51\textwidth}
\includegraphics[width=\textwidth]{../EGU2020/images/tweet.png}

{\footnotesize
Source: U.S. president via Twitter (2019-09-05)}
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Colors by designers, painters, and directors?}

\begin{minipage}[t]{0.65\textwidth}
\vspace*{-0.3cm}

\only<1>{\includegraphics[width=\textwidth]{images/almodovar-madre-original.png}}
\only<2>{\includegraphics[width=\textwidth]{images/almodovar-madre-palette.png}}
\only<3>{\includegraphics[width=\textwidth]{images/almodovar-madre-desaturated.png}}
\end{minipage}
\hfill
\begin{minipage}[t]{0.3\textwidth}

{\footnotesize

Movie: \emph{Todo sobre mi madre} (\emph{All About My Mother}, 1999)

\medskip

Picture source: Sony Pictures Classics via MoMA

\medskip

\only<2->{Palette: Hadley Mendelsohn}
}

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Colors by designers, painters, and directors?}

\vspace*{-0.3cm}

\begin{center}
\only<1>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-madre.png}}%
\only<2>{\includegraphics[width=.55\textwidth]{images/dorian-useR-2021-orrd.png}}%
\end{center}

\vspace*{-0.5em}
\begin{center}
{\footnotesize
Palette: \only<1>{Todo sobre mi madre}\only<2>{OrRd (ColorBrewer.org, HCL version)}\phantom{Org}}
\end{center}

\end{frame}

\begin{frame}
\frametitle{Colors by designers, painters, and directors?}

\begin{minipage}[t]{0.45\textwidth}
\vspace*{-0.3cm}

\only<1>{\includegraphics[width=\textwidth]{images/almodovar-tacones-original.png}}
\only<2>{\includegraphics[width=\textwidth]{images/almodovar-tacones-palette.png}}
\only<3>{\includegraphics[width=\textwidth]{images/almodovar-tacones-deutan.png}}
\end{minipage}
\hfill
\begin{minipage}[t]{0.5\textwidth}

{\footnotesize

Movie: \emph{Tacones lejanos} (\emph{High Heels}, 1991)

\medskip

Picture source: El Deseo S.A. via Twitter

\medskip

\only<2->{Palette: Bibiana Fernandez}
}

\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Wrap-up}

\textbf{Tools:}
\begin{itemize}
  \item \emph{grDevices:} \code{palette.colors()}, \code{hcl.colors()}.
  \item \emph{colorspace:} \code{swatchplot(..., cvd = TRUE)}.
  \item Interactive shiny apps on \url{https://www.hclwizard.org/}.
\end{itemize}

\pause
\bigskip

\textbf{Strategy:}
\begin{itemize}
  \item Check whether color is appropriate for coding your information.
  \item Use appropriate type of palette.
  \item Don't reinvent the wheel, start out from well-established palettes.
  \item Check robustness of palette.
  \item Be careful with palettes with too much chroma.
\end{itemize}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\small

Zeileis A, Fisher JC, Hornik K, Ihaka R, McWhite CD, Murrell P, Stauffer R, Wilke CO (2020).
  \dquote{colorspace: A Toolbox for Manipulating and Assessing Colors and Palettes.}
  \emph{Journal of Statistical Software}, \textbf{96}(1), 1--49.
  \doi{10.18637/jss.v096.i01}

\bigskip

Zeileis A, Murrell P, Maechler M, Sarkar D (2019).
  \dquote{A New \texttt{palette()} for R.}
  \emph{R~Foundation Blog}, 2019-11-21.
  \url{https://developer.R-project.org/Blog/public/2019/11/21/a-new-palette-for-r/}

\bigskip

Zeileis A, Murrell P (2019).
  \dquote{HCL-Based Color Palettes in grDevices.}
  \emph{R~Foundation Blog}, 2019-04-01.
  \url{https://developer.R-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/}

\bigskip
\bigskip

Web: \url{https://colorspace.R-Forge.R-project.org/}

\smallskip

Twitter: \href{https://twitter.com/AchimZeileis}{\tt @AchimZeileis}

\end{frame}

\end{document}